Accelerated Magnetic Resonance Imaging by Generative Adversarial Neural Networks

### 2018-02-27 letter.tex
Update

Dear Prof. XXX,

@@

It is our pleasure to submit a new manuscript entitled  ``{\it Accelerated Magnetic Resonance Imaging by Generative Adversarial Neural Networks}'' for
consideration by IEEE Transactions on Pattern Analysis and Machine Intelligence journal. 
The manuscript is co-authored by Ohad Shitrit and Tammy Riklin Raviv.

A preliminary version of this work was presented at the Deep Learning in Medical Image Analysis workshop (DLMIA 2017), Quebec, Canada
(in conjunction with the MICCAI conference). The submitted manuscript is substantially
longer and more comprehensive. We further extended the scope of our framework using a more stabilized training regime and exploiting the sparsity in two dimensions. The method is now tested on a large, diverse publicly available dataset of raw images without pre-processing, and with much higher sampling factors. Furthermore, extensive experiments were conducted to demonstrate the clinical applicability of the proposed framework and the strength of performing the reconstruction directly from the k-space.

Thank you for your consideration of this manuscript.

>>>>

Dear Prof. Dickinson,

@@

It is our pleasure to submit a new manuscript entitled  ``{\it Accelerated Magnetic Resonance Imaging by Generative Adversarial Neural Networks}'' for
consideration by the IEEE Transactions on Pattern Analysis and Machine Intelligence journal. 
The manuscript is co-authored by Ohad Shitrit and Tammy Riklin Raviv.

In this manuscript, we introduce a practical, software-only framework, based on deep learning, for accelerating MRI acquisition, while maintaining anatomically meaningful imaging. The key idea is performing MRI sub-sampling, while using a generative adversarial networks to directly estimate the missing k-space samples.
Comprehensive experiments using brain MRI scans, from a large publicly available dataset, demonstrate the quality of the proposed MRI reconstruction, obtained for up to six-fold acceleration factor. 
Our method is shown to outperform all the other tested methods, including the widely-used Compressed Sensing.

A preliminary version of this work was presented at the Deep Learning in Medical Image Analysis workshop (DLMIA 2017), Quebec, Canada
(in conjunction with the MICCAI conference). The submitted manuscript is substantially
longer and more comprehensive. We further extended the scope of our framework using a more stabilized training regime and exploiting the sparsity in two dimensions. The method is now tested on a large, diverse publicly available dataset of raw images without pre-processing, and with much higher sampling factors. Furthermore, extensive experiments were conducted to demonstrate the clinical applicability of the proposed framework and the advantage of performing the reconstruction directly from the k-space.

Thank you for the consideration of this manuscript.

### 2018-02-27 paper.tex
Update

\markboth{Journal of \LaTeX\ Class Files,~Vol.~XX, No.~XX, Feb~2018}%
{Shell \MakeLowercase{\textit{et al.}}: Accelerated Magnetic Resonance Imaging by Adversarial Neural Network}

@@

%--------------------- METHOD --------------------%
\section{Method}\label{method_section}
We aim to address MRI reconstruction from under-sampled k-space data using DNN. We hereby present the proposed methodology as well as the network architecture.
\subsection{Notations and Problem Definition}
Let $\bm{u}$ denote the desired signal, a 2D MRI, obtained by IFFT of the complex k-space signal $s_{0}$, i.e., 
\begin{equation}
\bm{u}=F^{H}s_{0}
\end{equation} 
where $H$ is the Hermitian transpose operation and $F^{H}$ is an orthonormal 2D IFFT operator, such that $F^{H}F=I$. 
Partial k-space sampling, results in MR images that are subject to artifacts and aliasing. 
Figure~\textbf{\ref{fig:pair_of_k_space_mri}}.d exemplifies artifacts caused by under-sampling and zero padding in the phase axis using a Gaussian mask (40\%).
\vspace{-0.5cm}
\begin{figure}[H]
	\centering
	\setlength\tabcolsep{1.5pt}
	\begin{tabular}{cccc}
		\subfloat[\textcolor{black}{\scriptsize{}}]{\centering{}\includegraphics[width=1.8cm,height=1.8cm]{include/grp1/example_mask_kspace_with_arrows}

@@

\subsection{System Architecture}
The proposed framework is illustrated in Figure~\ref{fig:system}.
\begin{figure}[H]
	\centering{}\includegraphics[width=8.0cm,height=4.5cm]{include/grp2/system_chart}\caption{\label{fig:system} \textcolor{black}{\footnotesize{}{}Framework
			architecture: $G$ and $D$ are the generator and discriminator networks,
			respectively. See text for details.}}
\end{figure}

@@

The training data consists of 500 3D brain MRI (T1-weighted) scans of different patients from the IXI dataset.\footnote{\url{http://brain-development.org/ixi-dataset/}} The data was acquired by three MR machines, Philips 1.5T, 3T and GE 1.5T using different imaging protocols.
From each 3D volume we extracted 93 2D saggital slices. All slices were interpolated to $256\times256$ pixels. We used $37.2k$ $(80\%)$ 2D slices for training and $9.3k$ $(20\%)$ for testing (100 3D volumes).
To create k-space images for training, inverse orthonormal
2D FFT was applied to the fully-sampled MR images. 
%Unlike our preliminary work~\cite{shitrit2017accelerated}, which employed a sampling mask along the phase axis alone; here, we perform k-space sampling using a 2D random Gaussian mask. By this, we exploit the sparsity in both phase and frequency dimensions and are therefore able to accommodate higher sampling factors (Figure~\ref{fig:sampling_masks}). MRI reconstruction is performed using 40\%, 25\%, and 16.6\% of the original k-space data.
We perform k-space sampling using a 2D random Gaussian mask. Thus, we exploit the sparsity in both phase and frequency dimensions and are therefore able to accommodate higher sampling factors (Figure~\ref{fig:sampling_masks}). MRI reconstruction is performed using 40\%, 25\%, and 16.6\% of the original k-space data.

@@

A key advantage of the proposed method is its short runtime at test phase (less than 10 msec per slice), which is almost two orders of magnitude faster than CS. This makes it applicable for routine uses. As training can be performed off-line, prior to the acquisition sessions, the fact that it lasts a few days has no practical relevance.

Assessing clinical usability of the reconstructed data is a considerable part in the quality analysis performed as part of this research. We chose to test it by using standard tissue segmentation measures (such as Dice and MHD), as we argue that compatibility of the extracted brain regions with those obtained by using the original fully-sampled MRIs indicate fine detail reconstructions that have apparent diagnostic significance. Nevertheless, pathological cases were not considered in this study and to the best of our knowledge are seldom addressed. We believe that high quality reconstruction in the presence of brain atrophies is the true success criterion. In fact, addressing this challenge and measuring performances in this sense, should be the gold standard in future benchmarks.

@@

%--------------------- Acknowledgment --------------------%
\subsection{Acknowledgment}
This study was partially supported by The Israel Science Foundation (1638/16); The Israel Defense Forces (IDF) Medical Corps and Directorate of Defense Research \& Development, Israeli Ministry of Defense (IMOD DDR\&D); Israel Ministry of Science, Technology and Space (63551).
%This research is partially supported by the Israel Science Foundation (T.R.R. 1638/16 ) and the IDF Medical Corps (T.R.R.)

@@

%-----------------------------BIBLIOGRAPHY--------------------%
\bibliography{paper}

>>>>

\markboth{IEEE TRANSACTIONS ON PATTERN ANALYSIS AND MACHINE INTELLIGENCE, ~VOL.~XX, NO.~XX, Feb~2018}%
{Shell \MakeLowercase{\textit{et al.}}: Accelerated Magnetic Resonance Imaging by Adversarial Neural Network}

@@

%--------------------- METHOD --------------------%
\section{Method}\label{method_section}
We aim to address MRI reconstruction from under-sampled k-space using DNN. We now present the proposed methodology and the network architecture.
\subsection{Notations and Problem Definition}
Let $\bm{u}$ denote the desired signal, a 2D MRI, obtained by IFFT of the complex k-space signal $s_{0}$, i.e., 
\begin{equation}
\bm{u}=F^{H}s_{0}
\end{equation} 
where $H$ is the Hermitian transpose operation and $F^{H}$ is an orthonormal 2D IFFT operator, such that $F^{H}F=I$. 
Partial k-space sampling, results in MR images that are subject to artifacts and aliasing. 
Figure~\textbf{\ref{fig:pair_of_k_space_mri}}.d exemplifies artifacts caused by under-sampling and zero padding in the phase axis using a Gaussian mask (40\%).
\vspace{-0.5cm}
\begin{figure}[H]
	\centering
	\setlength\tabcolsep{1.5pt}
	\begin{tabular}{cccc}
		\subfloat[\textcolor{black}{\scriptsize{}}]{\centering{}\includegraphics[width=1.8cm,height=1.8cm]{include/grp1/example_mask_kspace_with_arrows}

@@

\subsection{System Architecture}
The proposed framework is illustrated in Figure~\ref{fig:system}.
\begin{figure}[H]
	\centering{}\includegraphics[width=8.0cm,height=4cm]{include/grp2/system_chart}\caption{\label{fig:system} \textcolor{black}{\footnotesize{}{}Framework
			architecture: $G$ and $D$ are the generator and discriminator networks,
			respectively. See text for details.}}
\end{figure}

@@

The training data consists of 500 3D brain MRI (T1-weighted) scans of different patients from the IXI dataset.\footnote{\url{http://brain-development.org/ixi-dataset/}} The data was acquired by three MR machines, Philips 1.5T, 3T and GE 1.5T using different imaging protocols.
From each 3D volume we extracted 93 2D saggital slices. All slices were interpolated to $256\times256$ pixels. We used $37.2k$ $(80\%)$ 2D slices for training and $9.3k$ $(20\%)$ for testing (100 3D volumes).
To create k-space images for training, inverse orthonormal
2D FFT was applied to the fully-sampled MR images. 
%Unlike our preliminary work~\cite{shitrit2017accelerated}, which employed a sampling mask along the phase axis alone; here, we perform k-space sampling using a 2D random Gaussian mask. By this, we exploit the sparsity in both phase and frequency dimensions and are therefore able to accommodate higher sampling factors (Figure~\ref{fig:sampling_masks}). MRI reconstruction is performed using 40\%, 25\%, and 16.6\% of the original k-space data.
We perform k-space sampling using a 2D random Gaussian mask. Thus, we exploit the sparsity in both phase and frequency dimensions and are therefore able to accommodate higher sampling factors (Figure~\ref{fig:sampling_masks}). MRI reconstruction is performed using 40\%, 25\%, and 16.6\% of the original k-space.

@@

A key advantage of the proposed method is its short runtime at test phase (less than 10 msec per slice), which is almost two orders of magnitude faster than CS. This makes it applicable for routine uses. As training can be performed off-line, prior to the acquisition sessions, its runtime has no practical relevance.

%Assessing clinical usability of the reconstructed data is an important part in conducted experiments.
Quality analysis for assessment of the clinical usability of the reconstructed data plays a significant role in our experiments.
We chose to test it by using standard tissue segmentation measures (such as Dice and MHD), as we argue that compatibility of the extracted brain regions with those obtained by using the original fully-sampled MRIs indicate diagnostic significance. Nevertheless, pathological cases were not considered in this study and to the best of our knowledge are seldom addressed. We believe that high quality reconstruction in the presence of brain atrophies is the true success criterion. In fact, addressing this challenge and measuring performances in this sense, should be the gold standard in future benchmarks.

@@

%--------------------- Acknowledgment --------------------%
%\subsection{Acknowledgment}
\subsubsection*{\textbf{Acknowledgment}}
\small This study was partially supported by The Israel Science Foundation (1638/16); The Israel Defense Forces (IDF) Medical Corps and Directorate of Defense Research \& Development, Israeli Ministry of Defense (IMOD DDR\&D); Israel Ministry of Science, Technology and Space (63551).
%This research is partially supported by the Israel Science Foundation (T.R.R. 1638/16 ) and the IDF Medical Corps (T.R.R.)

@@

%-----------------------------BIBLIOGRAPHY--------------------%
\vspace{-0.3cm}
\bibliography{paper}
